---
title: "project"
author: "DawsonMacLean"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## CATAN

### Intro

Catan is a board game where players take on the role of settlers to obtain/trade resources and build roads/settlements. Resources include sheep, lumber, wheat, clay, and ore. A player's chance of winning the game is determined by both strategy and randomness (due to dice rolls). 

We got our data from Kaggle, and online community for data scientists. Our data contains information for 50 games with 4 players (200 entrees). The Kaggle data has information about the games such as dice rolls, number of resources gained/lost, turn order, and more. 

The goal of our project is to compare descriptive trends in the Catan data with what a randomforest model predicts are most important factors in who wins the game. Essentially, we are trying to find factors that predict a game win. 

Our "lie" for this project will be showing how descriptives (resources, game turn, piece color, etc.) contribute to who wins the game. Our graphs will make it appear as if certain descriptives help players to win more games compared to other descriptives.

In reality, the randomforest predictor model gives us a better overall view of what might be important predictors of game wins. This model contains a lot more data than our data set, so it is more likely to be less affected by outliers than our data set with only 50 games. 



## Modifying Data
One factor of board games that a lot of people are passionate about is the color of their game piece. It is an easy thing to get superstitious over as a result of attachment to their preferred color. The Catan data set did not have color listed as a variable, but we figured that people might be willing to interpret data about whether a color was more associated with wins, particularly if their color was favored. So, we started by manipulating the data set to include a variable for color. We set the seed and then randomly assigned colors to each of the four players in each game. As it was randomly sampled, with no connection to the original data, there are no *real* correlations to be found.

```{r Data Manipulation}
# Reading in Catan data from data folder
CatanData <- read.csv(file = 'data/catanstats.csv', header = TRUE)

set.seed(333)
# Creating object that has all of the desired colors
colors <- c("red", "blue", "white", "orange")

#applies all 4 colors per game with no repeats
CatanData$color <- unlist(
  lapply(
    split(CatanData, CatanData$gameNum),
    function(group)
    sample(colors))
  )



```


# We need this

Here we are gonna train a randomforest model thingy based on our 50 games. This allows us to compare descriptive information (graphs) with the randomforest model to determine predictors of winning

Dawson can write most/all of this ^^

```{r do_game,echo=TRUE,warning=FALSE,message=FALSE}
#,message=FALSE



#source("./simulatingCatan.R")

#to give the stats n plot for a Catan game, just enter a dataset of 4-player x 50 games (like the original)



#predictCatanPlayer(player, rfmodel)

```

# we need this 

Here are the "importance" values for the rf_model. To simplify, this refers to how important a variable is in predicting the likelihood of winning



  First we need to train the model on our data
  
  
```{r training, file = 'TrainRFModel.R'}


```

```{r train_now}
rf_model <- trainCatanModel(CatanData)

```


  Then we create functions to easily calculate the probability of winning (without going through all this every single time). One function is to predict for an entire set of 50 games, while the other is just for predicting based on one player's game stats.
```{r, file = 'simulatingCatan.R'} 

```

  The predicted probability of winning for every single player of every game (over 200 lines) is a lot, but the "importance" analysis of the model helps to understand what factors are relevant to predicting a win. 

```{r importance}
importance(rfmodel)

```

  Essentially

```{r importance}


#print(predictCatan)

#print(predictCatanPlayer)
#file='your-script.R'

```

## Descriptive Data Graphs


### Win rate by Color
Here are graphs showing descriptive characteritics of the oringal CatanData dataset.  

# *describe win by turn num*


```{r win_by_turn_num,echo=TRUE,warning=FALSE,message=FALSE}

#determining winner based on the first player to have >=10 points
CatanData$winloss <- rep("Loss", nrow(CatanData))

for (game in unique(CatanData$gameNum)) {
    
    maybe_win <- CatanData$player[CatanData$points >= 10 & CatanData$gameNum == game]
    
    if (length(maybe_win) > 0) {
      CatanData$winloss[((game - 1) * 4) + min(maybe_win)] <- "Win"
    }
  }

library(tidyverse)

library(ggplot2)

##win rate by player turn 

wins_only <- CatanData %>%
  filter(winloss == "Win")

win_rate_player <- CatanData %>%
  group_by(player) %>%
  summarise(
    total = n(),
    wins = sum(winloss == "Win"),
    win_rate = wins / total
  )

ggplot(win_rate_player, aes(x = factor(player), y = win_rate, fill = factor(player))) +
  geom_col() +
  labs(title = "Win Rate by Turn Position",
       x = "Turn Position (Player)",
       y = "Win Rate",
       fill = "Player") +
  scale_y_continuous(labels = scales::percent) +
  theme_minimal()

```

The second graph depicts the win rate by color. Based on the random data from the seed, more winners tend to pick blue, and the fewest winners tend to pick orange. Taking this data at face value, it seems that picking blue results in a significantly greater chance of winning a game of Catan than picking orange does. Those whose preferred color is blue might be inclined to believe these results, even though they are fully randomized within the 50 games of data and should have no real effect on game outcome.


```{r win_color}

##win rate by color 
win_rate_color <- CatanData %>%
  group_by(color) %>%
  summarise(
    total = n(),
    wins = sum(winloss == "Win"),
    win_rate = wins / total
  )

ggplot(win_rate_color, aes(x = color, y = win_rate, fill = color)) +
  geom_col() +  # Use 'geom_col' to create a bar plot
  scale_fill_manual(values = c("royalblue2", "darkgoldenrod1", "brown2", "snow2")) +  # Assign custom colors to each bar
  scale_y_continuous(labels = scales::percent_format()) +
  labs(title = "Win Rate by Color",
       x = "Color",
       y = "Win Rate") +
  theme_minimal()



```


### Win Rate by Starting Resource

  This graph highlights the impact of starting resources on winning likelihood. Wheat clearly stands out as the most winning-associated resource, followed by ore, sheep, wood (logs), and finally brick (clay). If you’re aiming for the best shot at victory, prioritizing wheat-heavy placements is a strong strategic move. The data also shows that starting with two of the same resource or building near the desert significantly decreases your chances — variety and access are key in early-game planning.



```{r win_resource}
##win rate resource
Catan_long <- CatanData %>%
  pivot_longer(
    cols = matches("settlenum|settlersc"),
    names_to = c("settlement", ".value"),
    names_pattern = "(\\dsettle)(num\\d|rsc\\d)"
  ) %>%
  rename(number = num1, resource = rsc1)  # fix names for consistency

Catan_long <- Catan_long %>%
  mutate(
    number = as.numeric(number),
    resource = str_trim(resource),
    resource = ifelse(str_detect(resource, "^[A-Z]$"), resource, NA)
  ) %>%
  drop_na(number, resource)

win_rate_resource <- Catan_long %>%
  group_by(resource) %>%
  summarise(
    total = n(),
    wins = sum(winloss == "Win"),
    win_rate = wins / total
  )

label_map <- c("L" = "Lumber", "C" = "Clay", "S" = "Sheep", "W" = "Wheat", "O" = "Ore")

# Generate the plot
ggplot(win_rate_resource, aes(x = resource, y = win_rate, fill = resource)) +
  geom_col() +
  scale_fill_manual(values = c("L" = "forestgreen",    # Lumber (L) = Green
                               "C" = "chocolate",   # Clay (C) = Brown
                               "S" = "darkolivegreen2",       # Sheep (S) = Blue
                               "W" = "goldenrod1",     # Wheat (W) = Yellow
                               "O" = "cadetblue4")) +       # Ore (O) = Gray
  scale_y_continuous(labels = scales::percent_format()) +
  scale_x_discrete(labels = label_map) +  # Replace abbreviations with full names
  labs(title = "Win Rate by Starting Resource",
       x = "Resource",
       y = "Win Rate") +
  theme_minimal()









```

# we nee conclusion

## Conclusions

### Conclusions based on Graphs

  player 2 is the best, I think wheat is the best not sure  
  
  just describe what the plots would suggest is important
  
  
  
### Conclusions based on RandomForest

  probaly will put up the importance chart thing again
  
  explain what it probably means
  
  
  
### Differences

  explain what is different quickly
  
  
  probably will put in a short predictCatanPlayer() thing where we will only differ the color, to show how that compares with the gra  phs
  
  any difference with color being significant should mean nothing
  
  
## yippee we done

short goodbe conclusion
  
  

