---
title: "project"
author: "DawsonMacLean"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## CATAN

### Intro

  what is Catan, where we got our data

  goal: compare descriptive trends in the Catan data with what a randomforest model predicts what are most important factors

  "LIE" : using just descriptive

  "TRUTH": using randomforest model as a more reliable way to show what is actually important. While also knowing that it is incomplete


## Modifying Data
One factor of board games that a lot of people are passionate about is the color of their game piece. It is an easy thing to get superstitious over as a result of attachment to their preferred color. The Catan data set did not have color listed as a variable, but we figured that people might be willing to interpret data about whether a color was more associated with wins, particularly if their color was favored. So, we started by manipulating the data set to include a variable for color. We set the seed and then randomly assigned colors to each of the four players in each game. 

```{r Data Manipulation}
# Reading in Catan data from data folder
CatanData <- read.csv(file = 'data/catanstats.csv', header = TRUE)

set.seed(333)
# Creating object that has all of the desired colors
colors <- c("red", "blue", "white", "orange")

#applies all 4 colors per game with no repeats
CatanData$color <- unlist(
  lapply(
    split(CatanData, CatanData$gameNum),
    function(group)
    sample(colors))
  )



```




Here we are gonna train a randomforest model thingy based on our 50 games. This allows us to compare descriptive information (graphs) with the randomforest model to determine predictors of winning

Dawson can write most/all of this ^^

```{r do_game,echo=TRUE,warning=FALSE,message=FALSE}
#,message=FALSE



source("./simulatingCatan.R")

#to give the stats n plot for a Catan game, just enter a dataset of 4-player x 50 games (like the original)


predictCatan(CatanData,rf_model)

#predictCatanPlayer(player, rfmodel)

```

Here are the "importance" values for the rf_model. To simplify, this refers to how important a variable is in predicting the likelihood of winning

Dawson can write most/all of this ^^

```{r importance}

print(importance(rf_model))

print(trainCatanModel)

print(predictCatan)

print(predictCatanPlayer)


```

## Descriptive Data Graphs

Here are graphs showing descriptive characteritics of the oringal CatanData dataset.  

The second graph depicts the win rate by color. Based on the random data from the seed, more winners tend to pick blue, and the fewest winners tend to pick orange. Taking this data at face value, it seems that picking blue results in a significantly greater chance of winning a game of Catan than picking orange does. Those whose preferred color is blue might be inclined to believe these results, even though they are fully randomized within the 50 games of data and should have no real effect on game outcome.

```{r graphs,echo=TRUE,warning=FALSE,message=FALSE}

#determining winner based on the first player to have >=10 points
CatanData$winloss <- rep("Loss", nrow(CatanData))

for (game in unique(CatanData$gameNum)) {
    
    maybe_win <- CatanData$player[CatanData$points >= 10 & CatanData$gameNum == game]
    
    if (length(maybe_win) > 0) {
      CatanData$winloss[((game - 1) * 4) + min(maybe_win)] <- "Win"
    }
  }

library(tidyverse)

library(ggplot2)

##win rate by player turn 

wins_only <- CatanData %>%
  filter(winloss == "Win")

win_rate_player <- CatanData %>%
  group_by(player) %>%
  summarise(
    total = n(),
    wins = sum(winloss == "Win"),
    win_rate = wins / total
  )

ggplot(win_rate_player, aes(x = factor(player), y = win_rate, fill = factor(player))) +
  geom_col() +
  labs(title = "Win Rate by Turn Position",
       x = "Turn Position (Player)",
       y = "Win Rate",
       fill = "Player") +
  scale_y_continuous(labels = scales::percent) +
  theme_minimal()

##win rate by color 
win_rate_color <- CatanData %>%
  group_by(color) %>%
  summarise(
    total = n(),
    wins = sum(winloss == "Win"),
    win_rate = wins / total
  )

ggplot(win_rate_color, aes(x = color, y = win_rate, fill = color)) +
  geom_col() +  # Use 'geom_col' to create a bar plot
  scale_fill_manual(values = c("royalblue2", "darkgoldenrod1", "brown2", "snow2")) +  # Assign custom colors to each bar
  scale_y_continuous(labels = scales::percent_format()) +
  labs(title = "Win Rate by Color",
       x = "Color",
       y = "Win Rate") +
  theme_minimal()

##win rate resource
Catan_long <- CatanData %>%
  pivot_longer(
    cols = matches("settlenum|settlersc"),
    names_to = c("settlement", ".value"),
    names_pattern = "(\\dsettle)(num\\d|rsc\\d)"
  ) %>%
  rename(number = num1, resource = rsc1)  # fix names for consistency

Catan_long <- Catan_long %>%
  mutate(
    number = as.numeric(number),
    resource = str_trim(resource),
    resource = ifelse(str_detect(resource, "^[A-Z]$"), resource, NA)
  ) %>%
  drop_na(number, resource)

win_rate_resource <- Catan_long %>%
  group_by(resource) %>%
  summarise(
    total = n(),
    wins = sum(winloss == "Win"),
    win_rate = wins / total
  )

label_map <- c("L" = "Lumber", "C" = "Clay", "S" = "Sheep", "W" = "Wheat", "O" = "Ore")

# Generate the plot
ggplot(win_rate_resource, aes(x = resource, y = win_rate, fill = resource)) +
  geom_col() +
  scale_fill_manual(values = c("L" = "forestgreen",    # Lumber (L) = Green
                               "C" = "chocolate",   # Clay (C) = Brown
                               "S" = "darkolivegreen2",       # Sheep (S) = Blue
                               "W" = "goldenrod1",     # Wheat (W) = Yellow
                               "O" = "cadetblue4")) +       # Ore (O) = Gray
  scale_y_continuous(labels = scales::percent_format()) +
  scale_x_discrete(labels = label_map) +  # Replace abbreviations with full names
  labs(title = "Win Rate by Starting Resource",
       x = "Resource",
       y = "Win Rate") +
  theme_minimal()









```

## Bootstrap?

do we source the bootstrap file or is that kinda too much stuff at this point


## Conclusions

### Conclusions based on Graphs

  player 2 is the best, I think wheat is the best not sure  
  
  just describe what the plots would suggest is important
  
  
  
### Conclusions based on RandomForest

  probaly will put up the importance chart thing again
  
  explain what it probably means
  
  
  
### Differences

  explain what is different quickly
  
  
  probably will put in a short predictCatanPlayer() thing where we will only differ the color, to show how that compares with the graphs
  
  any difference with color being significant should mean nothing
  
  
## yippee we done

short goodbe conclusion
  
  

